The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-19 paper/mri_paper.tex
knobloch and julien, pessah

The MRI evolves via a period of linear growth, followed by nonlinear saturation. The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive. Most previous works on MRI saturation have studied the problem in simulation. Indeed, only a few works have tackled MRI saturation analytically. We briefly summarize a few of these analytic contributions.

@@

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime. The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.

@@

>>>>

The MRI evolves via a period of linear growth, followed by nonlinear saturation. The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive. Most previous works on MRI saturation have studied the problem in simulation. Indeed, only a few works have tackled MRI saturation analytically. \citeauthor{Knobloch:2005ba} (\citeyear{Knobloch:2005ba}) analyze the MRI in the strongly nonlinear regime, by following the already-developed MRI modes into asymptotic saturation. They find that this saturation proceeds via modification of the background shear. Another notable analytic work is that of \citeauthor{Pessah:2010ic} (\citeyear{Pessah:2010ic}), who presents an analysis of MRI saturation via secondary instabilities. This work suggests that MRI saturation can be caused by parasite instabilities growing off of the primary MRI modes, in particular by Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.

@@

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime. The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space. In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.

@@

\bibitem[{Knobloch \& Julien(2005)}]{Knobloch:2005ba}
Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106

\bibitem[Pessah(2010)]{Pessah:2010ic}
Pessah, M, 2010, ApJ, 716, 1012
